---
title: "Day 7 Practice"
output: html_document
---

In this practice, let's explore data on tipping, in the `tips` data frame from the `reshape2` package.

0. First get a little familiar with the structure of the data, using `summary()`, `head()`, etc.  What are the numeric variables?  What are their units? What about factors? Are factor levels sensible?

```{r}
library(reshape2)
library(ggplot2)

mytips <- tips
head(tips)
summary(tips)
str(tips)

#Numeric Variables: total_bill & tip; their units are dollors

#Factor Variables: sex, smoker, day, time; the levels are sensible


```

1. Create a simple scatterplot, plotting the tip amount by the total bill.
  - first plot the raw quantities
  - "save" the initial simple plot as an object
  - change the axis labels for both x and y
    - try two different methods
  
```{r}
plot1a <- ggplot(mytips, aes(total_bill, tip))+geom_point()
plot1a

plot1b <- ggplot(mytips, aes(total_bill, tip))+geom_point()
         +xlab("Tip Amount")+ylab("Total Bill")
plot1b

plot1c <- ggplot(mytips, aes(total_bill, tip))+geom_point()+scale_x_continuous(name = "Tip Amount")+scale_y_continuous(name = "Total Bill")
plot1c


```

2. Continue using the previous data
  - using `scales`, plot the axes so that they follow a logarithmic pattern
  - add "break" values every 5 dollars on the x axis, and every 2 dollars on the y axis
  - finally, use `labels = dollar` (made available by the `scales` package) to format the axis numbers as dollars

```{r}
library(scales)
plot2.log <- ggplot(mytips, aes(total_bill, tip))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10()
 
plot2.log

plot2.break <- ggplot(mytips, aes(total_bill, tip))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(breaks = seq(0, 50, 5)) + scale_y_continuous(breaks = seq(0, 10, 2))
 
plot2.break


plot2.dollar <- ggplot(mytips, aes(total_bill, tip))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar, breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar, breaks = seq(0, 10, 2))
 
plot2.dollar

```

3. Continue to build off of the final scatterplot in the previous problem (hint: you may need to change the plot object, in order to make the following changes)
  - map gender to the color of the points
  - change the color values, using both the "brewer" and "manual" methods
  - add regression lines (regular linear regression), which also break down by gender
  - also try a loess smoother instead of regression lines
  - finally, change the overall "theme" to either the "classic", "minimal", or "bw" theme

```{r}
plot3.sex <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))

plot3.sex

plot3.sex.brewer <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_brewer(palette = "Dark2")

plot3.sex.brewer

plot3.sex.manual <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(breaks = seq(0, 50, 5)) + scale_y_continuous(breaks = seq(0, 10, 2))+ scale_color_manual(values = c("pink1", "skyblue1"))

plot3.sex.manual


plot3.sex.regression <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_manual(values = c("pink1", "skyblue1"))+geom_smooth(method = "lm")

plot3.sex.regression

plot3.sex.loess <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(breaks = seq(0, 50, 5)) + scale_y_continuous(breaks = seq(0, 10, 2))+ scale_color_manual(values = c("pink1", "skyblue1"))+geom_smooth()  + theme_bw()

plot3.sex.loess

plot3.sex.classic <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_manual(values = c("pink1", "skyblue1"))+geom_smooth(method = lm)  + theme_classic()

plot3.sex.classic

```


4. Now let's modify the legend for the `sex` variable, building off off the scatterplot with linear regression lines
  - using the appropriate `scale` function, change the name of the legend to "Gender"
  - change the location of the legend with the `theme()` function
  - also using the "legend.position" argument, remove the legend completely
  - finally, put the legend back to the default position, but change the levels, such that Male is given as the first level
  
```{r}
plot4 <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_manual(name = "Gender", values = c("pink1", "skyblue1"))+geom_smooth(method = lm)  + theme_classic() +theme(legend.position = "top")

plot4

mytips$sex
mytips$sex <- factor(mytips$sex, levels = rev(levels(mytips$sex)))
mytips$sex

plot4a <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_manual(name = "Gender", values = c("pink1", "skyblue1"))+geom_smooth(method = lm)  + theme_classic() + theme(text = element_text(family = "serif", size = 12))

plot4a
```

5. Let's explore how this pattern might vary for different days of the week.
  - use faceting to create a separate plot for each day
  - try both the "wrap" and "grid" faceting functions
  - make a change so that the labels on the top of the facets give the full day name (e.g., "Friday", not "Fri"), and they are ordered from Thursday to Sunday
  
```{r}

levels(mytips$day)
levels(mytips$day) <- c("Friday", "Saturday", "Sunday", "Thursday")
mytips$day = factor(mytips$day, levels = c("Thursday", "Friday", "Saturday", "Sunday"))
levels(mytips$day)

plot5 <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_manual(name = "Gender", values = c("pink1", "skyblue1"))+geom_smooth(method = lm)  + theme_bw() +theme(legend.position = "top") + facet_wrap(~day)



plot5

plot5a <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_manual(name = "Gender", values = c("pink1", "skyblue1"))+geom_smooth(method = lm)  + theme_bw() +theme(legend.position = "top") + facet_grid(.~day)

plot5a

plot5b <- ggplot(mytips, aes(total_bill, tip, color = sex))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_manual(name = "Gender", values = c("pink1", "skyblue1"))+geom_smooth(method = lm)  + theme_bw() +theme(legend.position = "top") + facet_grid(day~.)

plot5b

```

6. Now let's look at adding information about the size of the party to the plot
  - first leave off the regression lines, to make it easier to see
  - map party size to the size of the points
  - look at both an aggregate plot and plots faceted by day
  - now add back the regression lines, but only plot a single regression line (i.e., not broken down by gender), remove the confidence interval (hint: look at the arguments for `geom_smooth()`), and make the line a dashed black line
  
```{r}

plot6 <- ggplot(mytips, aes(total_bill, tip, color = sex, size=size))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_manual(name = "Gender", values = c("pink1", "skyblue1")) +geom_smooth(method = lm, se=FALSE,linetype = "dashed", color = "black")+ theme_bw() +theme(legend.position = "top") 

plot6

plot6a <- ggplot(mytips, aes(total_bill, tip, color = sex, size=size))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_manual(name = "Gender", values = c("pink1", "skyblue1")) +geom_smooth(method = lm, se=FALSE,linetype = "dashed", color = "black")+ theme_bw()  + facet_wrap(~day)

plot6a

plot6b <- ggplot(mytips, aes(total_bill, tip, color = sex, size=size))+geom_point() + xlab("Tip Amount") + ylab("Total Bill") + scale_x_log10(labels = dollar,breaks = seq(0, 50, 5)) + scale_y_continuous(labels = dollar,breaks = seq(0, 10, 2))+ scale_color_manual(name = "Gender", values = c("pink1", "skyblue1")) +geom_smooth(method = lm, se=FALSE,linetype = "dashed", color = "black")+ theme_bw()  + facet_grid(day~.)

plot6b

```

