---
title: "PSYC 798W Final Project"
author: "Jason Hasbrouck"
date: "1/18/2018"
output: html_document
---
For my final project I will use the terrorism data set to answer the following questions: 

1) Is there a difference in frequency of attacks between years?
   - 1a)Is there a difference in severity of attacks between years?

2) Is there a difference in frequency of attacks based on month of year? 
   - 2a)Is there a difference in severity of attacks based on month of year?

3) Which region of the world has largest volume of terror attacks?

4) What types of attacks lead to the most fatalities?

5) What types of attacks lead to the most injuries?

For questions 1 and 2, I will run a correlation test to determine the relationships between frequency in attacks based on months and on years.Then I will use bootstrapping by creating my own loops in order to compare the results from the cor.test and the results from bootstrapping for the 95% confidence intervals. 

Then for  questions 3 - 5, I will use either multiple regressions,  ANOVAs, or BAYES Factors to answer the questions. I will provide plots to display my findings for each question. 

```{r}
library(dplyr)
library(ggplot2)
library(plyr)
library(tidyr)
library(tidyverse)
library(reshape2)
library(psych)
library(boot)
library(BayesFactor)
library(readxl)
library(RColorBrewer)

set.seed(22)

terrorism <- read_excel("~/Library/Mobile Documents/com~apple~CloudDocs/Desktop/R Homework/R_Homework_Jason_Hasbrouck/terrorism.xlsx")
my.terrorism <- terrorism

summary(is.na(my.terrorism))

summary(my.terrorism)

#Removing rows with mission data -- most missing data was in the columns containing information on number of fatalities (nkill) and number of wounded (nwound)
my.terrorism <- na.omit(my.terrorism)

summary(is.na(my.terrorism))


my.terrorism$iyear <- as.factor(my.terrorism$iyear)
my.terrorism$imonth <- as.numeric(my.terrorism$imonth)
my.terrorism$attacktype1 <- as.numeric(my.terrorism$attacktype1)
my.terrorism$attacktype1_txt <- as.factor(my.terrorism$attacktype1_txt)

my.terrorism$weaptype1_txt <- as.factor(my.terrorism$weaptype1_txt)


my.terrorism$attacktype1b <- my.terrorism$attacktype1
levels(my.terrorism$attacktype1b) <- c("assassination", "armed.assault", "bombing.explosion", "hijacking", "hostage.barricade", "hostage.kidnapping", "facility.attack", "unarmed.assault", "unknown" )
my.terrorism$attacktype1b <- as.factor(my.terrorism$attacktype1b)
my.terrorism$attack.ID <- 1:nrow(my.terrorism)
my.terrorism$casualty <- my.terrorism$nkill + my.terrorism$nwound
my.terrorism$casualty <- as.numeric(my.terrorism$casualty)

```

#1: Is there a difference in frequency of attacks between years?
##  - 1a) Is there a difference in severity of attacks between years?

### 1 The results show that the most terror attacks occured in 2014 (16860), followed by 2015 (14825), 2016 (13488), and 2013 (11996).

### 1a) The results  show a significant differerence in severity of attacks for the following years: 

- 2013 & 2014: t = 2.6974, df = 24734, p-value = 0.006992

- 2013 & 2015: t = -2.0415, df = 24627, p-value = 0.04122

- 2013 & 2016: t = -2.3923, df = 20688, p-value = 0.01675

- 2014 & 2015: t = -4.8378, df = 27336, p-value = 1.32e-06

- 2014 & 2016: t = -4.4859, df = 19821, p-value = 7.301e-06
 
```{r}

#Number of Attacks Annualy
summary(my.terrorism$iyear)


#Graph of the number of Terrorist Attacks Annually
frequency.annual.attacks <- ggplot(my.terrorism, aes(iyear)) + geom_bar(stat = "count")+ylab("Number of Attacks")+xlab("Year") +scale_y_continuous(breaks = c(2000, 4000, 6000, 8000, 10000, 12000, 14000, 16000)) + ggtitle("Annual Number of Terrorist Attacks Worldwide")+ theme(plot.title = element_text(hjust = 0.5))

frequency.annual.attacks


#Comparing relationships between number of attacks and year
attack.2013 <- droplevels(my.terrorism[my.terrorism$iyear == "2013",  ])

attack.2014 <- droplevels(my.terrorism[my.terrorism$iyear == "2014",  ])

attack.2015 <- droplevels(my.terrorism[my.terrorism$iyear == "2015",  ])

attack.2016 <- droplevels(my.terrorism[my.terrorism$iyear == "2016",  ])


t.test(attack.2013$casualty, attack.2014$casualty)

t.test(attack.2013$casualty, attack.2015$casualty)

t.test(attack.2013$casualty, attack.2016$casualty)

t.test(attack.2014$casualty, attack.2015$casualty)

t.test(attack.2014$casualty, attack.2016$casualty)

t.test(attack.2015$casualty, attack.2016$casualty)

casualty.per.attack.annual <- ggplot(my.terrorism, aes(iyear, casualty))+geom_boxplot(aes(fill = iyear)) +scale_y_log10(breaks = c(0, 2,4,6,8,100,500, 1000,1500))+ylab("Casualty per Attack")+ scale_fill_brewer(palette = "Set1") +xlab("Year")+ggtitle("Casualties Per Terrorist Attack")+ theme(plot.title = element_text(hjust = 0.5), legend.position = "none")

casualty.per.attack.annual

#Set up Wide Data Frame:

   

#Setting Up Bootstrap: 

annual <- function(formula, data, indices){
  d <- data[indices,]
  fit <- lm(formula, data = d)
  return(summary(fit)$r.square)
}

monthly.cor.results <- boot(data=month.attack.wide, statistic=monthly.rsq, R=200, formula = ID ~ January+February+March+April+May+June+July+August+September+October+November+December)
plot(monthly.cor.results)

monthly.bootcorr <- boot(month.attack.wide, monthly.cor, R=500)
summary(monthly.bootcorr)

plot(monthly.bootcorr)

monthly.boot.ci <- boot.ci(monthly.bootcorr, type="perc")
```

#2 Is there a difference in frequency of attacks based on month of year? 
  - The month with the most number of terror attacks between 2013 and 2016 was May. 

May: 5426
Jul: 5261
Oct: 4880
Jan: 4876
Aug: 4821
Apr: 4791
Nov: 4697
Mar: 4636
Feb: 4511
Jun: 4564
Sep: 4376
Dec: 4357


  - 1a) Is there a difference in severity of attacks between years?
      - A Spearman test for correlation shows that theree is no linear relationship between severity of terrorist attack and month `R^2` = .007 (t = 1.658, df = 51955, p-value = 0.09732; 95% Confidence Interval = 1.156868 to 1.444911).

  - Bootstrapping Results: 


```{r}
#Graph of the number of Terrorist Attacks By Month
frequency.monthly.attacks <- ggplot(my.terrorism, aes(imonth)) + geom_bar(stat = "count")+ylab("Number of Attacks")+xlab("Year") + scale_x_continuous(name = "Month", breaks = 1:12, labels = month.name)+scale_y_continuous(breaks = c(1000, 2000, 3000, 4000, 5000)) +ggtitle("Monthly Number of Terrorist Attacks (2013 - 2016")+ theme(plot.title = element_text(hjust = 0.5))

frequency.monthly.attacks


#Comparing relationships between number of attacks and year
jan.attack <- droplevels(my.terrorism[my.terrorism$imonth == "1",  ])
count(jan.attack$imonth)

feb.attack <- droplevels(my.terrorism[my.terrorism$imonth == "2",  ])
count(feb.attack$imonth)

mar.attack <- droplevels(my.terrorism[my.terrorism$imonth == "3",  ])
count(mar.attack$imonth)

apr.attack <- droplevels(my.terrorism[my.terrorism$imonth == "4",  ])
count(apr.attack$imonth)

may.attack <- droplevels(my.terrorism[my.terrorism$imonth == "5",  ])
count(may.attack$imonth)

jun.attack <- droplevels(my.terrorism[my.terrorism$imonth == "6",  ])
count(jun.attack$imonth)

jul.attack <- droplevels(my.terrorism[my.terrorism$imonth == "7",  ])
count(jul.attack$imonth)

aug.attack <- droplevels(my.terrorism[my.terrorism$imonth == "8",  ])
count(aug.attack$imonth)

sep.attack <- droplevels(my.terrorism[my.terrorism$imonth == "9",  ])
count(sep.attack$imonth)

oct.attack <- droplevels(my.terrorism[my.terrorism$imonth == "10",  ])
count(oct.attack$imonth)

nov.attack <- droplevels(my.terrorism[my.terrorism$imonth == "11",  ])
count(nov.attack$imonth)

dec.attack <- droplevels(my.terrorism[my.terrorism$imonth == "12",  ])
count(dec.attack$imonth)

# cor.test comparing if there is a difference between month and number of casualties
cor.test(my.terrorism$imonth, my.terrorism$casualty) 



#Graph of Monthly Casualties in Terror Attacks (2013 - 2016)
casualty.per.attack.monthly <- ggplot(my.terrorism, aes(imonth, log(casualty)))+geom_bar(stat = "identity", aes(fill = iyear))+ scale_fill_brewer(palette = "Set1", name = "Year")+ scale_x_continuous(name = "Month", breaks = 1:12, labels = month.name)+xlab("Month")+ylab("Number of Casualties")+ggtitle("Monthly Casualties in Terror Attacks (2013 - 2016)")+ theme(plot.title = element_text(hjust = 0.5))

casualty.per.attack.monthly



#Create Monthly Data in Wide Format
month.attack<- data.frame(my.terrorism$imonth, my.terrorism$casualty)
head(month.attack)

colnames(month.attack) <- c("month", "casualty")
head(month.attack)

month.attack$ID <- 1:nrow(month.attack)
str(month.attack)


month.attack.wide <- spread(month.attack, month, casualty, fill = 0)

colnames(month.attack.wide) <- c("ID", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

str(month.attack.wide)


#Run correlation for each month

month.cor <- month.attack.wide[,2:13]
round(cor(month.cor), 2)

#Set Up  Bootstrap for Correlation CI:

monthly.rsq <- function(formula, data, indices){
  d <- data[indices,]
  fit <- lm(formula, data = d)
  return(summary(fit)$r.square)
}

monthly.cor.results <- boot(data=month.attack.wide, statistic=monthly.rsq, R=200, formula = ID ~ January+February+March+April+May+June+July+August+September+October+November+December)
plot(monthly.cor.results)

monthly.bootcorr <- boot(month.attack.wide, monthly.cor, R=500)
summary(monthly.bootcorr)

plot(monthly.bootcorr)

monthly.boot.ci <- boot.ci(monthly.bootcorr, type="perc")


```

#3 Which region of the world has largest volume of terror attacks?

## The following list represtents the number of terror attacks by region from 2013 - 2016
   1:   23509 - Middle East & North Africa 
   2:   17805 - South Asia 
   3:   7287 - Sub-Saharan Africa  
   4:   4416 - Southeast Asia 
   5:   1946 - Eastern Europe 
   6:   1073 - Western Europe 
   7:   769 - South America   
   8:   201 - North America   
   9:   92 - East Asia    
   10:  41 - Central Asia
   11:  34 - Australasia & Oceania    
   12:  23 - Central America & Caribbean    

```{r}

#Volume of Terror Attacks by Region

my.terrorism$region_txt <- as.factor(my.terrorism$region_txt)
summary(my.terrorism$region_txt)

count(my.terrorism$region_txt)



#Graph of the volume of Terrorist Attacks by Region 
region.attack.volume <- ggplot(my.terrorism, aes(region_txt)) + geom_bar(stat = "count")+ylab("Number of Attacks")+xlab("Region") +scale_y_log10(breaks = c(1000, 5000,10000, 15000, 20000))+ ggtitle("Volume of Terrorist Attacks by Region (2013 - 2016)")+ theme(plot.title = element_text(hjust = 0.5)) 

region.attack.volume

```

#4 What types of attacks lead to the most fatalities?

##Hostage Barricade incidents are the terrorist attack type that typically lead to the highest number of fatalities

###Additionally, a Bayes Factor analysis shows that overall the type of attack is a moderate predictor of number of Fatalities (BF10 = 3.64)

### A Bayesian Regression shows very strong support that the type of terror attacks that leads to the highest number of fatalities is Hostage taking via kidnapping (BF10 = 46.784)

List of Bayes Factors by Attack Type: 
Unarmed Assault     3.435
Facility Attack     0.012
Hostage Kidnapping  46.784
Hostage Barricade   0.0249
Hijacking           0.015
Bombing Explosion   13.904
Armed Assault       0.013
Assassination       0.012
Unknown             2.203

```{r}
#Assumption Checking

#1 Linearity - categorical variables do have a linear relationship, but there are significant outliers -- Assumption Met
ggplot(my.terrorism, aes(attacktype1_txt, nkill)) +geom_point()

#2 Normality of Residuals - The data has an extreme positive skew. Will correct with a square root transformation. -- Assumption Met
lm.fatality <- lm(nkill ~ attacktype1_txt, data = my.terrorism)

ggplot(lm.fatality, aes(lm.fatality$residuals))+geom_histogram(binwidth = 5)
ggplot(lm.fatality, aes(log(lm.fatality$residuals)))+geom_histogram(binwidth = 1)



#3 Homogenaity of Residuals: - Unable to plot due to error 
## -- keep getting error: Error: Aesthetics must be either length 1 or the same as the data (57196): x, y
qplot(lm.fatality$residuals, geom = "histogram")
summary(lm.fatality)

#4 Continuous, Interval level data -- Assumption Met

#5 Data Points are independent -- Assumption Met



#Graph of Fatalities by Terror Attack Type
ggplot(my.terrorism, aes(attacktype1_txt, log(nkill), fill = attacktype1b))+geom_boxplot()+
  ylab("Number of Fatalities")+
  xlab("Attack Type") + 
  scale_fill_brewer(palette = "Paired", name = "Attack Type") +
  ggtitle("Fatalities by Terror Attack Type")+ 
  theme(plot.title = element_text(hjust = 0.5))



#General Bayes Test of all attack types 
BF.fatality <- generalTestBF(formula=nkill~attacktype1_txt, as.data.frame(my.terrorism))
BF.fatality

```
#Bayesian Analysis: Attack type by injury
- Will be hidden from knitted file due to size output
```{r eval = FALSE, echo = FALSE}
#Setting Up Data for Bayes Test by Attack Type 
attack.kill <- data.frame(my.terrorism$attacktype1b, my.terrorism$nkill)

colnames(attack.kill) <- c("attacktype", "nkill")

attack.kill$ID <- 1:nrow(attack.kill)

head(attack.kill)

attack.kill.wide <- spread(attack.kill, attacktype, nkill, fill = 0)
head(attack.kill.wide)

str(attack.kill.wide)

#General Bayes Test of Each Attack Type by casualty
BF.attack.kill<- generalTestBF(ID~assassination+armed.assault+bombing.explosion+hijacking+hostage.barricade+hostage.kidnapping 
+facility.attack+unarmed.assault+unknown,data=attack.kill.wide)
BF.attack.kill
```


#5 What types of attacks lead to the most injuries?

##Hostage Barricade incidents are the terrorist attack type that typically lead to the highest number of injuries.

###Additionally, a Bayes Factor analysis shows that overall the type of attack is a moderate predictor of number of Injuries (BF10 = 9.20)

### A Bayesian Regression shows very little evidence that individual attack type is a strong predictor of number of injuries associated with a terror attack.

List of Bayes Factors by Attack Type: 
Unarmed Assault     0.010
Facility Attack     0.026
Hostage Kidnapping  0.023
Hostage Barricade   0.026
Hijacking           0.010
Bombing Explosion   0.010
Armed Assault       0.043
Assassination       0.215
Unknown             2.203

```{r}

#Assumption Checking

#1 Linearity - categorical variables do have a linear relationship, but there are significant outliers -- Assumption Met
ggplot(my.terrorism, aes(attacktype1_txt, nwound)) +geom_point()

#2 Normality of Residuals - The data has an extreme positive skew. Will correct with a square root transformation. -- Assumption Met
lm.injury <- lm(nwound ~ attacktype1_txt, data = my.terrorism)

ggplot(lm.injury, aes(lm.injury$residuals))+geom_histogram(binwidth = 5)
ggplot(lm.injury, aes(log(lm.injury$residuals)))+geom_histogram(binwidth = 1)



#3 Homogenaity of Residuals: - Unable to plot due to error 
## -- keep getting error: Error: Aesthetics must be either length 1 or the same as the data (57196): x, y
qplot(lm.injury$residuals, geom = "histogram")
summary(lm.injury)

#4 Continuous, Interval level data -- Assumption Met

#5 Data Points are independent -- Assumption Met


BF.injury <- generalTestBF(formula=nwound~attacktype1_txt, as.data.frame(my.terrorism))
BF.injury

ggplot(my.terrorism, aes(attacktype1_txt, log(nwound), fill = attacktype1b))+geom_boxplot()+
  ylab("Number of Injuries")+
  xlab("Attack Type") + 
  scale_fill_brewer(palette = "Paired", name = "Attack Type") +
  ggtitle("Injuries by Attack Type")+ 
  theme(plot.title = element_text(hjust = 0.5))


#Setting Up Data for Bayes Test by Attack Type 
attack.wound <- data.frame(my.terrorism$attacktype1b, my.terrorism$nwound)

colnames(attack.wound) <- c("attacktype", "nwound")

attack.wound$ID <- 1:nrow(attack.wound)

head(attack.wound)

attack.wound.wide <- spread(attack.wound, attacktype, nwound, fill = 0)
head(attack.wound.wide)

```
#Bayesian Analysis: Attack type by injury
- Will be hidden from knitted file due to size output
```{r eval = FALSE, echo = FALSE}
BF.attack.wound<- generalTestBF(formula=ID~assassination + armed.assault + bombing.explosion + hijacking+ hostage.barricade+hostage.kidnapping 
 + facility.attack  + unarmed.assault, as.data.frame(attack.wound.wide))
BF.attack.wound
```
     
