---
title: "PSYC 798W Final Project"
author: "Jason Hasbrouck"
date: "1/18/2018"
output: html_document
---

For my final project I will use the terrorism data set to answer the following questions: 

1) Is there a difference in severity of attacks between years?

2) Is there a difference in severity of attacks based on month of year?

3) Which region of the world has largest volume of terror attacks?

4) What types of attacks lead to the most fatalities?

5) What types of attacks lead to the most injuries?

For questions 1 and 2, I will run a correlation test to determine the relationships between severity in attacks based on months and on years.Then I will use bootstrapping by creating my own loops in order to compare the results from the cor.test and the results from bootstrapping for the 95% confidence intervals. 

Then for  questions 3 - 5, I will use either multiple regressions,  ANOVAs, or BAYES Factors to answer the questions. I will provide plots to display my findings for each question. 


```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(reshape2)
library(psych)
library(BayesFactor)
library(readxl)
library(RColorBrewer)

terrorism <- read_excel("terrorism.xlsx")
my.terrorism <- terrorism

my.terrorism$iyear <- as.factor(my.terrorism$iyear)
my.terrorism$imonth <- as.numeric(my.terrorism$imonth)
my.terrorism$attacktype1 <- as.numeric(my.terrorism$attacktype1)
my.terrorism$attacktype1_txt <- as.factor(my.terrorism$attacktype1_txt)

my.terrorism$weaptype1_txt <- as.factor(my.terrorism$weaptype1_txt)


my.terrorism$attacktype1b <- my.terrorism$attacktype1
levels(my.terrorism$attacktype1b) <- c("assassination", "armed.assault", "bombing.explosion", "hijacking", "hostage.barricade", "hostage.kidnapping", "facility.attack", "unarmed.assault", "unknown" )
my.terrorism$attacktype1b <- as.factor(my.terrorism$attacktype1b)
my.terrorism$attack.ID <- 1:nrow(my.terrorism)
my.terrorism$casualty <- my.terrorism$nkill + my.terrorism$nwound
my.terrorism$casualty <- as.numeric(my.terrorism$casualty)


?spread
ID.wide <- spread(my.terrorism, iyear, attack.ID, convert = TRUE)

attack.means <- summarize(group_by(my.terrorism, attacktype1b, iyear), mean = mean(casualty, na.rm = TRUE))
attack.means

```

#1: Is there a difference in severity of attacks between years?

The results  show a significant differerence in severity of attacks for the following years: 

- 2013 & 2014: t = 2.6974, df = 24734, p-value = 0.006992

- 2013 & 2015: t = -2.0415, df = 24627, p-value = 0.04122

- 2013 & 2016: t = -2.3923, df = 20688, p-value = 0.01675

- 2014 & 2015: t = -4.8378, df = 27336, p-value = 1.32e-06

- 2014 & 2016: t = -4.4859, df = 19821, p-value = 7.301e-06
 

```{r}

attack.2013 <- droplevels(my.terrorism[my.terrorism$iyear == "2013",  ])

attack.2014 <- droplevels(my.terrorism[my.terrorism$iyear == "2014",  ])

attack.2015 <- droplevels(my.terrorism[my.terrorism$iyear == "2015",  ])

attack.2016 <- droplevels(my.terrorism[my.terrorism$iyear == "2016",  ])


t.test(attack.2013$casualty, attack.2014$casualty)

t.test(attack.2013$casualty, attack.2015$casualty)

t.test(attack.2013$casualty, attack.2016$casualty)

t.test(attack.2014$casualty, attack.2015$casualty)

t.test(attack.2014$casualty, attack.2016$casualty)

t.test(attack.2015$casualty, attack.2016$casualty)

ggplot(my.terrorism, aes(iyear, casualty))+geom_boxplot() +scale_y_log10(breaks = c(0, 2,4,6,8,100,500, 1000,1500))+ylab("Casualty per Attack")+xlab("Year")+ggtitle("Casualties Per Terrorist Attack")+ theme(plot.title = element_text(hjust = 0.5))
```


#2 Is there a difference in severity of attacks based on month of year?

```{r}

```


#3 Which region of the world has largest volume of terror attacks?

```{r}

```

#4 What types of attacks lead to the most fatalities?

```{r}

```

#5 What types of attacks lead to the most injuries?

```{r}

plot(my.terrorism$nkill ~ my.terrorism$attacktype1_txt)

lm.injury <- lm(nkill ~ attacktype1_txt, data = my.terrorism)

summary(lm.injury)


```

