---
title: "Day 8 Homework"
output: html_document
---

Find two repetitive things that you would like to be able to do with your data.  This can be something having to do with coding variables, or saving plots, or running an analysis multiple times on different subsets, etc. etc.  Just try to find something where you would normally need to "copy and paste" sections of code, varying only one or two things.  Figure out how to use a loop to accomplish this.  

For BOTH of these things, take the following steps:

1. Articulate what you want to do (verbally)

2. Give some code that illustrates the "copy and paste" method.  This can just show a couple of steps.  That is, if you would be running through a bunch of iterations, just copy and paste a couple of times, not everything.

3. Work this code into a loop format, so that no copy & paste is neccesary.  You will need to set up a "counter" variable that cycles through various values (which do not have to be numeric), changing with every pass through the loop.  But you may also need to make other changes to your code to accommodate the loop.

As usual, send me your .Rmd file with any data I need to replicate your results.

#Data
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(reshape2)
library(readxl)
library(RColorBrewer)

terrorism <- read_excel("terrorism.xlsx")
my.terrorism <- terrorism

my.terrorism$iyear <- as.factor(my.terrorism$iyear)
my.terrorism$imonth <- as.factor(my.terrorism$imonth)
my.terrorism$attacktype1 <- as.factor(my.terrorism$attacktype1)
my.terrorism$attacktype1_txt <- as.factor(my.terrorism$attacktype1_txt)
my.terrorism$weaptype1_txt <- as.factor(my.terrorism$weaptype1_txt)
my.terrorism$attacktype1b <- my.terrorism$attacktype1
levels(my.terrorism$attacktype1b) <- c("assassination", "armed.assault", "bombing.explosion", "hijacking", "hostage.barricade", "hostage.kidnapping", "facility.attack", "unarmed.assault", "unknown" )
my.terrorism$attacktype1b <- as.factor(my.terrorism$attacktype1b)
my.terrorism$attack.ID <- 1:nrow(my.terrorism)
my.terrorism$casualty <- my.terrorism$nkill + my.terrorism$nwound
my.terrorism$casualty <- as.numeric(my.terrorism$casualty)
```

#LOOP 1:

##Create New Column that indicates the level of terrorist attack severity (low, medium, high)

  - To do this I will create a loop that will create a new column and then code the variables as low, medium, and high depending on how many casualties there were. 
  
```{r}
#COPY AND PASTE CODE EXAMPLE:

non.loop <- my.terrorism
non.loop$severity <- non.loop$casualty
non.loop$severity <- ifelse(non.loop$severity < 5, "low", NA)

#and so on for "medium" and "high"


#LOOP CODE EXAMPLE

loop1 <- my.terrorism

loop1$attack.severity <- ifelse(loop1$casualty < 5, "low", ifelse(loop1$casualty > 20, "high",ifelse(loop1$casualty >= 5 & loop1$casualty <=20, "medium", NA)))

```

#LOOP 2

## Create a loop of plots for severity of attacks by each year


```{r}
levels(my.terrorism$weaptype1_txt) <- c("Biological", "Chemical", "Explosives", "Fake Weapons", "Firearms", "Incendiary", "Melee", "Other", "Sabotage Equipment", "Unknown", "Vehicle")
levels(my.terrorism$weaptype1_txt)

ggplot(my.terrorism, aes(x = iyear, y = casualty, fill = weaptype1_txt))+geom_bar(stat = "identity", position = "dodge") +scale_y_log10(breaks = c(10, 50, 250, 500, 750, 1000, 1500)) +ggtitle("Annual Terrorist Attack Casualties by Weapon Type") +scale_fill_brewer(palette = "Spectral", name = "Weapon Type")+ theme(plot.title = element_text(hjust = 0.5))

# Would then continue building above ggplot for each type of weapon individually, instead, I could use the following loop: 

pdf("loop.plots.homework8.pdf")
for(this.weapon in levels(my.terrorism$weaptype1_txt)) {
print(ggplot(my.terrorism[my.terrorism$weaptype1_txt == this.weapon, ], aes(iyear,casualty)) + geom_bar(stat = "identity") + ggtitle(this.weapon))
}
dev.off()


```

