---
title: "Day 10 lecture/review"
output: html_document
---

# Template for working with new functions

## Some steps
  1. Find the function/package you want
    - Google
    - People you know
    - TaskViews: R home page > link on the left for Packages > link for CRAN Task Views
    - Papers (contact authors if necessary)
  2. If you find a package:
    - Look at the CRAN page
      - Google: "CRAN <packagename>"
      - website links (URL)
      - vignettes
  3. When you find a function
    1. go to the help page (`?<functionname>`)
    2. what are the required arguments?
      - in Usage
      - what arguments have defaults? what are those defaults?
    3. what kinds of things can go in the required arguments?
      - type of object: look in Arguments
      - format
      - `example()`
    4. try it!
    5. inspect the structure of the resulting object
    6. troubleshoot
    
## Example: factor analysis

1. Google around: decide on `factanal()` (also `principal()` from the `psych` package, or the various functions in the `lavaan()` package)
2. Already in `stats` package
3. `?factanal`
4. Required arguments:
  - `x`
  - `factors`
  
```{r}
myris <- iris
myris.fa1 <- factanal(x = ~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width, data = myris, factors = 1)
myris.fa2 <- factanal(x = myris[, c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")], factors = 1)

myris.fa1 # same as print(myris.fa1)
myris.fa2
names(myris.fa1)
str(myris.fa1)
class(myris.fa1)

# help() : Value
```

  - reminder: 
    - ERRORS mean "I give up, there's a problem"
    - WARNINGS mean "you might not want to do that (but I did it anyway)"
  - inspect the structure of the resulting object
    - `print()` (same as just typing name of object + Enter)
    - `names()`
    - `str()`
    - look at the Value part of the help file
  - browse through optional arguments


# Template for loops (or vectorization)

Example:
  - We want a regression
  - estimate, std err, p-value
  - scatterplot with regression line
  - for each species separately

1. DO IT ONCE!
  - do whatever you want to do, on just one chunk of your data (or whatever it is that you want to iterate through)

```{r}
library(ggplot2)
myris <- iris

myris.subset <- myris[myris$Species == "setosa", ] # or filter(myris, Species == "setosa")

#run regression
myris.lm <- lm(Sepal.Width ~ Sepal.Length, data = myris.subset) 

#pull out values
coefs <- summary(myris.lm)$coefficients
est <- coefs["Sepal.Length", "Estimate"]
std.error <- coefs["Sepal.Length", "Std. Error"]
pval <- coefs["Sepal.Length", "Pr(>|t|)"]

#build a plot
ggplot(myris.subset, aes(Sepal.Length, Sepal.Width)) + geom_point() + geom_smooth(method = "lm")

```

2. Figure out what MINIMALLY needs to change for each pass through a loop
  - try a copy & paste, get it to work for the next thing
  - the species
  - THAT minimal thing is what you want the counter to loop through (or index)
  
```{r}
myris <- iris


#Species type "virginica" is the one thing that we want to change for our loop
myris.subset <- myris[myris$Species == "virginica", ] # or filter(myris, Species == "virginica")

myris.lm <- lm(Sepal.Width ~ Sepal.Length, data = myris.subset) 
coefs <- summary(myris.lm)$coefficients
est <- coefs["Sepal.Length", "Estimate"]
std.error <- coefs["Sepal.Length", "Std. Error"]
pval <- coefs["Sepal.Length", "Pr(>|t|)"]

ggplot(myris.subset, aes(Sepal.Length, Sepal.Width)) + geom_point() + geom_smooth(method = "lm")

```

3. Modify to fit the loop structure
  - figure out how to get all the counter values in a vector
  - `for(COUNTER in COUNTER.VECTOR) { }`
  - set up "results storage" object
    - list
    - data.frame
    - etc.
    - check to make sure it looks like what you want
  - how do I get the values produced in each loop?
    - think about the merits of string vector "counter" vs. a numeric one
    
```{r}
#for(this.species in c("setosa", "virginica", "versicolor")) {
#for(this.species in levels(myris$Species)) {
#for(this.species in unique(myris$Species)) {
#
all.species <- levels(myris$Species)

# storing results
# my.results <- list(est = NA, std.errors = NA, pvals = NA, plots = NA) # list also possible, just maybe less convenient

my.results <- data.frame(species = all.species, est = NA, std.err = NA, pval = NA)
my.results # check this "empty" results table 
my.plots <- list()
for(this.species in all.species) {
  print(paste("Now we're getting the results for", this.species))
  myris.subset <- myris[myris$Species == this.species, ]
  myris.lm <- lm(Sepal.Width ~ Sepal.Length, data = myris.subset) 
  coefs <- summary(myris.lm)$coefficients
  
  # the values we want
  est <- coefs["Sepal.Length", "Estimate"]
  std.error <- coefs["Sepal.Length", "Std. Error"]
  pval <- coefs["Sepal.Length", "Pr(>|t|)"]
  plot <- ggplot(myris.subset, aes(Sepal.Length, Sepal.Width)) + geom_point() + geom_smooth(method = "lm") + labs(title = this.species)
  
  # stick the values in our results object
  my.results[my.results$species == this.species, c("est", "std.err", "pval")] <- c(est, std.error, pval)
  my.plots[[this.species]] <- plot
}

my.results
my.plots[["setosa"]]
my.plots[["virginica"]]
my.plots[["versicolor"]]

#this.species <- "setosa"
#  print(ggplot(myris.subset, aes(Sepal.Length, Sepal.Width)) + geom_point() + geom_smooth(method = "lm") + labs(title = this.species))

# faceting: sometimes a good alternative to looping (with plots)

ggplot(myris, aes(Sepal.Length, Sepal.Width)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~ Species)

library(grid)
library(gridExtra)
grid.arrange(my.plots[[1]], my.plots[[2]], my.plots[[3]], ncol = 1) # good for arranging multiple plots in one window/page

```

NOTE: sometimes "indices" work better than "values"

```{r}
all.species <- levels(myris$Species)

my.results <- data.frame(species = all.species, est = NA, std.err = NA, pval = NA)
my.plots <- list()

for(this.row in 1:length(all.species)) {
  print(paste("Now we're getting the results for", all.species[this.row]))
  myris.subset <- myris[myris$Species == all.species[this.row], ]
  myris.lm <- lm(Sepal.Width ~ Sepal.Length, data = myris.subset) 
  coefs <- summary(myris.lm)$coefficients
  
  # the values we want
  est <- coefs["Sepal.Length", "Estimate"]
  std.error <- coefs["Sepal.Length", "Std. Error"]
  pval <- coefs["Sepal.Length", "Pr(>|t|)"]
  plot <- ggplot(myris.subset, aes(Sepal.Length, Sepal.Width)) + geom_point() + geom_smooth(method = "lm") + labs(title = all.species[this.row])
  
  # stick the values in our results object
  my.results[this.row, c("est", "std.err", "pval")] <- c(est, std.error, pval)
  my.plots[[this.row]] <- plot
}

my.results
names(my.plots)
str(my.plots)
my.plots[[1]]
my.plots[[2]]
my.plots[[3]]
```

4. When things break: 
  - find out where it broke (if an Error)
    - last value of counter == where it broke
  - run in through single loops
    - manually set counter variable

## Vectorized version
  - same overall process
    1. DO IT ONCE
    2. Instead of finding counter variable, find arguments
       - write down (comments) what you expect each argument to be

```{r}

get.results <- function(chunk, DV, IV) {
  # expect chunk to be a data frame
  # expect DV and IV to be strings, names of columns to go in the regression
  my.formula <- as.formula(paste(DV, "~", IV)) # handy way to "build" a formula
  myris.lm <- lm(my.formula, data = chunk) 
  #my.iris.lm <- lm(get(DV))~data = chunk)
  coefs <- summary(myris.lm)$coefficients
  
  # the values we want
  est <- coefs[IV, "Estimate"]
  std.error <- coefs[IV, "Std. Error"]
  pval <- coefs[IV, "Pr(>|t|)"]
  # plot <- ggplot(chunk, aes(get(IV), get(DV))) + geom_point() + geom_smooth(method = "lm") + labs(title = this.species)
  
  # stick the values in our results object
  my.results <- data.frame(est = est, std.err = std.error, pval = pval)

  return(my.results)
}

get.results(myris, DV = "Petal.Width", IV = "Petal.Length")

library(dplyr)

# with pipes
all.results <- myris %>% group_by(Species) %>% do(get.results(chunk = ., DV = "Petal.Width", IV = "Petal.Length")) %>% as.data.frame()
all.results

# without pipes
all.results2 <- as.data.frame(do(.data = group_by(myris, Species), get.results(.,  DV = "Petal.Width", IV = "Petal.Length")))
all.results2


all.results

```

# Debugging functions

- by-hand debugging

```{r}
# "manually" set the value of each argument, step through the function as if it were just a regular script
IV <- "Petal.Width"
DV <- "Petal.Length"
chunk <- myris
```

- with `debug()` and `browser()`

```{r}
test.function <- function(chunk, DV, IV) {
  my.formula <- as.formula(paste(DV, "~", IV))
  myris.lm <- lm(my.formula, data = chunk) 
  coefs <- summary(myris.lm)$coefficients
  
  # the values we want
  est <- coefs[IV, "Estimate"]
  std.err <- coefs[IV, "Std. Error"]
  pval <- coefs[IV, "Pr(>t|)"]
  plot <- ggplot(chunk, aes(get(IV), get(DV))) + geom_point() + geom_smooth(method = "lm") + labs(title = this.species)
  
  # stick the values in our results object
  my.results <- data.frame(est = est, std.err = std.err, pval = pval)

  return(my.results)
  
}

test.function(myris, "Sepal.Width", "Sepal.Length")

debug(test.function)

test.function(myris, "Sepal.Width", "Sepal.Length")

# capital "Q" + Enter exits debugger (aka "browser")

# can also insert browser()
test.function2 <- function(chunk, DV, IV) {
  my.formula <- as.formula(paste(DV, "~", IV))
  myris.lm <- lm(my.formula, data = chunk) 
  coefs <- summary(myris.lm)$coefficients
  #if(this.species == "setosa") { # if you want to selectively browse
    browser()
  #}
 
  # the values we want
  est <- coefs[IV, "Estimate"]
  std.err <- coefs[IV, "Std. Error"]
  pval <- coefs[IV, "Pr(>t|)"]
  plot <- ggplot(chunk, aes(get(IV), get(DV))) + geom_point() + geom_smooth(method = "lm") + labs(title = this.species)
   # stick the values in our results object
  my.results <- data.frame(est = est, std.err = std.err, pval = pval)
  return(my.results)
}

test.function2(myris, "Sepal.Width", "Sepal.Length")

```

# More examples

## EXAMPLE: SD-based cutoffs

```{r}
my.iris <- iris

#if you want to break up your data among one dimension, you want to get data into "long" format so you can use dplyr

my.iris$plantID <- 1:nrow(my.iris)

mean.value <- mean(my.iris$Sepal.Length)
sd.value <- sd(my.iris$Sepal.Length)
min.value <- mean.value - sd.value *1
max.value <- mean.value + sd.value*1

#want to arrange data so that if data is below the min value, we make it the min value and do the same for values above max value

## to do this use ifelse

my.iris$Sepal.Length.Truncated <- ifelse(my.iris$Sepal.Length < min.value, yes = min.value, ifelse(my.iris$Sepal.Length > max.value, yes = max.value, no = my.iris$Sepal.Length))


#check that values changed
summary(my.iris)


#creating function
truncate.by.sd <- function(chunk, value.column, multiplier = 1) {
  chunk <- as.data.frame(chunk)
  mean.value <- mean(chunk[, value.column])
  sd.value <- sd(chunk[, value.column])
  min.value <- mean.value - sd.value * multiplier
  max.value <- mean.value + sd.value* multiplier
  chunk$Sepal.Length.Truncated <- ifelse(chunk[, value.column] < min.value, yes = min.value, ifelse(chunk[, value.column] > max.value, yes = max.value, no = chunk[, value.column]))
  return(chunk)
}

my.iris2 <- truncate.by.sd(my.iris, "Sepal.Length")

library(dplyr)
my.iris3 <- group_by(my.iris, Species) %>% do(truncate.by.sd (., "Sepal.Length"))

summary(my.iris3)

library(tidyr)
my.iris.long <- gather(my.iris, variable, value, Sepal.Length:Petal.Width)

head(my.iris.long)

my.iris4 <- group_by(my.iris.long, Species, variable) %>% do(truncate.by.sd(.,"value"))

head(my.iris4)


my.iris4.wide <- spread(my.iris4, variable, value)

summary(my.iris4.wide)

```

## EXAMPLE: Error bars

```{r}
head(my.iris.long)

#want to get cell means by species and variable

std.err <- function(x) {
  se <- sd(x)/sqrt(length(x))
  se
}

cell.means <- group_by(my.iris.long, variable, Species) %>% summarize(mean.value = mean(value), se = std.err(value))

cell.means 


ggplot(cell.means, aes(variable, mean.value))+
  geom_bar(stat="identity", position = "dodge", aes(fill = "Species")) +geom_errorbar(ymin = min.value - 1.96*se, ymax = mean.value+1.96*se, position = "dodge")


```


#When to use vectorizing functions: 

Scott's rules of thumb: 
  - If you are doing something to each row in your data or to each column ... you can use `apply()`
  - If you want to do some function to each member of a vector or a list (one-dimensional), you can use `sapply` or `lapply()` if you want to get a list back
  - EVERYWHERE else try to use the group_by() function in dplyr
    - may require you to set up a table of conditions
    - sometimes summarize()
      - want to boil down a lot of numbers (or some other vector) to one
    - sometimes mutate()
      - based on transformation of another column
      - with rowwise() instead of group_by(), you get the same effect as apply()
    - sometimes do()
      - anything more complicated, write a function and use do()
      - you take a chunk of a data frame and return a data frame
      - 
