---
title: "Day 5 Homework"
output: html_document
---

0. Reload your data, and re-use the code you already have for getting things formatted nicely (columns renamed, factors relabeled, etc.)

```{r}
library(dplyr)
library(tidyr)
library(readxl)
terrorism <- read_excel("~/Desktop/R Homework/R_Homework_Jason_Hasbrouck/terrorism.xlsx")

my.terrorism <- terrorism


my.terrorism$iyear <- as.factor(my.terrorism$iyear)

my.terrorism$imonth <- as.factor(my.terrorism$imonth)

my.terrorism$attacktype1 <- as.factor(my.terrorism$attacktype1)

my.terrorism$attacktype1_txt <- as.factor(my.terrorism$attacktype1_txt)

str(my.terrorism)

my.terrorism$attacktype1b <- my.terrorism$attacktype1

levels(my.terrorism$attacktype1b) <- c("assassination", "armed.assault", "bombing.explosion", "hijacking", "hostage.barricade", "hostage.kidnapping", "facility.attack", "unarmed.assault", "unknown" )

levels(my.terrorism$attacktype1b)

my.terrorism$attacktype1b <- as.factor(my.terrorism$attacktype1b)

str(my.terrorism)
```

1. Aggregate (e.g., get mean/median/max/something) of some numerical variable, over one or more factors.  Use `group_by()` plus `summarize()` from the `dplyr` package. You may or may not need to reshape the data first!

```{r}
my.terrorism$attack.ID <- 1:nrow(my.terrorism)

my.terrorism$casualty <- my.terrorism$nkill + my.terrorism$nwound

attack.means <- summarize(group_by(my.terrorism, attacktype1b, iyear), mean = mean(casualty, na.rm = TRUE))
attack.means

```

2. Run TWO DIFFERENT simple analyses or statistical tests, such as linear regression (`lm()`), logistic regression (`glm()`), correlation test (`cor.test()`), t-test (`t.test()`), or non-parametric tests (e.g., `wilcox.test()`).  For each of these:
  - Describe why you are doing this analysis, i.e., what question is it answering?
  - I won't judge you on statistical expertise!  (though I will make comments if I think I can be helpful)
  - Report some key statistics from the analysis, using **inline** code. HINT: outside of an R code chunk, use `r codehere` format.
  
```{r}



```

3. Get your data into "long" (aka "tidy") format, if it isn't already

```{r}
#Data is currently in long format
```







