---
title: "Day 5 Practice"
output: html_document
---

1. Make another copy of the iris data set
  - Using `data.frame()` create a (small) table that assigns a color to each species.  That is, make a column called Species and a column called Color, with one row per species
  - use `merge()` to merge the data into the copy of the iris data set
  - create another table with an additional species (and color)
  - merge this new table with iris. What happened?
  - create one more table matching species to color, with only two rows, missing setosa
  - merge this new table with the copy of iris. What happened?

```{r}

library(dplyr)
iris.practice <- iris

iris.color <- data.frame(Species = c("setosa", "versicolor", "virginica"), color = c("red", "purple", "pink"))


iris.merge <- merge(iris.practice, iris.color)


iris.color.extra <- data.frame(Species = c("setosa", "versicolor", "virginica", "extra"), color = c("red", "purple", "pink", "orange"))

iris.color.nosetosa <- data.frame(Species = c("versicolor", "virginica"), color = c("purple", "pink"))

iris.merge.extra <- merge(iris.practice, iris.color.extra, all = FALSE)
iris.merge.extra2 <- merge(iris.practice, iris.color.extra, all = TRUE)
iris.merge.extra3 <- left_join(iris.practice, iris.color.extra)
iris.merge.extra3$Species <- as.factor(iris.merge.extra3$Species)

nrow(iris.practice)
nrow(iris.merge.extra)
nrow(iris.merge.extra2)
nrow(iris.merge.extra3)

#if you gain or lose a lot of rows, use xtabs to see what happened

iris.merge.nosetosa <- merge(iris.color, iris.color.nosetosa)

```

2. Make (yet another) copy of the iris data set
  - Create a column that takes the log of the Sepal.Width
  - Create a column that represents the multiplication of Sepal.Width by Sepal.Length
  - Create a "median split" column on Sepal.Width. In other words, calculate the median of Sepal.Width, and then create a column that indicates whether the value in that row is less than the median or not
  
```{r}
iris.practice <- iris
iris.practice$Sepal.Width.log <- log(iris.practice$Sepal.Width)
head(iris.practice)

iris.practice$length.by.width <- iris.practice$Sepal.Length * iris.practice$Sepal.Width
head(iris.practice)

median(iris.practice$Sepal.Width)

iris.practice$median.split <- iris.practice$Sepal.Width < 3
head(iris.practice)

#another method
iris.practice$median.split[iris.practice$Sepal.Width < 3] <- "low"
iris.practice$median.split[iris.practice$Sepal.Width >= 3] <- "high"

#another method
iris.practice$median.split2 <- ifelse(test = iris.practice$Sepal.Width<3, yes = "low", no = "high")

iris.practice$median.split3 <- as.factor(ifelse(test = iris.practice$Sepal.Width<3, yes = "low", no = "high"))

xtabs(~ Sepal.Width + median.split, iris.practice)
xtabs(~ Sepal.Width + median.split2, iris.practice)
xtabs(~ Sepal.Width + median.split3, iris.practice)


```

3. Start with (yes, you guessed it!) another copy of the iris data set
  - Use `gather()` to convert this wide data to long format
    - You should be "gathering up" all of the columns with measurements
  - "Tidy" up the new factor column by splitting it into two columns: one designating dimension (width vs. length), one designating flower part (sepal vs. petal)
    
```{r}

my.iris <- iris

#Create a column for Plant ID 
my.iris$PlantID <- 1:nrow(my.iris)

my.iris.long <- gather(my.iris, "measurement", "centimeters", Sepal.Length : Petal.Width) 
head(my.iris.long)

#seperate(data, column, where to separate, what to name new columns)   use"\\." to ensure you split along the dot. if you just use sep = "." it will try to split on every character
my.iris.long <- separate(my.iris.long, measurement, sep = "\\.", c("plant.part", "dimension"))
head(my.iris.long)
summary(my.iris.long)

```

4. Continue working with the "long" iris data from previous problem
  - using `dplyr` functions, get the mean for each "cell" in the species x part x dimension design
  - make the results "wide", with different columns for each species

```{r}

```

5. Use the sample data in the scottdata folder
  - run a logistic regression, using `glm()`, where the "Merged" variable is the (binary) dependent variable, and "WithinMinPairs" is the predictor.
  - Hint: the basic usage of `glm()` is:

results <- glm(DV ~ predictor1 + predictor2 + ...  ,                               
               data = dataframe.object, family = "binomial")
               
```{r}
practice <- read.csv("scottdata/CognitionPaperFinalData.csv")

```

  - use `summary()` to display the results.
  
```{r}

```

  - using INLINE code, report the estimate, standard error, and p-value for the WithinMinPairs coefficient.  Hint #1: save the summary as an object, and pull values from that. Hint #2: use `str()` to figure out how to extract these values. Hint #3: save the individual results as objects first, then put in inline code.

```{r}

```

The estimate is `r`, the standard error is `r`, and the p-value is `r`.

EXTRA CREDIT: figure out how to display the above results with only a few decimal places

